\documentclass[10pt]{article}

\usepackage{amssymb, amsfonts, amsmath, amsthm}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[margin=1.00 in]{geometry}
\usepackage{enumerate}
\usepackage{harpoon}
\usepackage{nicefrac}
\usepackage{tikz, pgfplots}
\usepackage{xypic}
\usepackage{tikz-cd}
\usepackage{booktabs}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{} % clear all header fields
\renewcommand{\headrulewidth}{0pt} % no line in header area
\fancyfoot{} % clear all footer fields
\fancyfoot[LE,RO]{\footnotesize \thepage}           % page number in "outer" position of footer line

\pgfplotsset{compat=newest}

%\setlength\parindent{0pt}%removes indents from entire file

\newtheorem*{sol}{Solution}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\Unif}{\operatorname{Uniform}}
\newcommand{\Bern}{\operatorname{Bernoulli}}
\newcommand{\Binom}{\operatorname{Binomial}}
\newcommand{\Poiss}{\operatorname{Poisson}}
\newcommand{\se}{\operatorname{se}}

\begin{document}

\noindent \large{Solutions to selected exercises from Chapter 10 of
\emph{Wasserman --- All of Statistics}}

\begin{enumerate}
\item[(5)] Let $X_1,\ldots,X_n \sim \Unif(0,\theta)$ and let
$Y = \max\{X_1,\ldots,X_n\}$ We want to test
\[H_0 : \theta = 1/2 \text{ versus } H_1 : \theta > 1/2.\]
The Wald test is not appropriate since $Y$ does not converge to a Normal.
Suppose we decide to test this hypothesis by rejecting $H_0$ when $Y>c$.
\begin{enumerate}[(a)]
\item Find the power function.

\begin{sol}
Choose some $c \geq 0$. The rejection region is
\[
R = \left\{ (x_1,\ldots,x_n) : x_i > c \text{ for some } i\right\}.
\]

Hence the power function is $\beta(\theta)=\P_\theta((X_1,\ldots,X_n)\in R)$
and this is
\[
\P_\theta(X_i>c \text{ for some } i) = 1 - \P_\theta(X_i<c \text{ for all } i)
= 1 - \Pi_{i=1}^n \P_\theta(X_i < c).
\]
If $c\geq\theta$ then $\P_\theta(X_i < c) = 1$ for all $i$. If $c<\theta$ then
$\P_\theta(X_i < c) = c/\theta$ for all $i$. Hence
\[
\beta(\theta) =
\begin{cases}
1-(c/\theta)^n & \theta > c \\
0 & \theta \leq c\\
\end{cases}
\]
\end{sol}
\item What choice of $c$ will make the size of the test $.05$?
\begin{sol}
The size is
\[
\alpha = \sup_{\theta=1/2} \beta(\theta) = \beta(1/2) = 1-(2c)^n.
\]
So we need to solve $1-(2c)^n=0.05$, i.e. $(2c)^n = 19/20$. The solution
is
\[
c=\frac{1}{2}\left(\frac{19}{20}\right)^{1/n}.
\]
\end{sol}
\item In a sample of size $n=20$ with $Y=0.48$ what is the p-value? What
conclusion about $H_0$ would you make?
\begin{sol}
Choosing the parameter $c$, the size is $\alpha=1-(2c)^n$ with corresponding
rejection region
$R_\alpha = \left\{ (x_1,\ldots,x_n) : x_i > c \text{ for some } i\right\}$.
The p-value is
\[
\inf \{\alpha \in (0,1) : Y \in R_\alpha\} = \inf\{\alpha \in (0,1) : Y > c\}
\]
Since $\alpha$ is strictly monotonically decreasing with $c$ the p-value
is exactly the value of $\alpha$ when $Y=c$. In other words, the p-value is
\[\alpha=1-(2Y)^n=1-(2 \cdot 0.48)^{20} \approx 0.558.\]
We would not reject the null hypothesis $H_0$ in this case.
\end{sol}
\item In a sample of size $n=20$ with $Y=0.52$ what is the p-value? What
conclusion about $H_0$ would you make?
\begin{sol}
Of course we can reject $H_0$ in this case: $\theta$ is definitely
not equal to $1/2$. Let's confirm this using the p-value.
The p-value in this case is
\[
\inf\{\alpha \in (0,1) : Y > c\} =
\inf \{\alpha \in (0,1) : Y > 1/2(1-\alpha)^{1/20}\}.
\]
This is equal to 0 since $Y > 1/2$
and we reject $H_0$ since our p-value $0$ is $<0.05$.
\end{sol}
\end{enumerate}

\item[(6)] There is a theory that people can postpone their death until after an
important event. To test the theory, Phillips and King (1988) collected
data on deaths around the Jewish holiday Passover. Of 1919 deaths, 922
died the week before the holiday and 997 died the week after. Think of
this as a binomial and test the null hypothesis that $\theta=1/2$. Report and
interpret the p-value. Also construct a confidence interval for $\theta$.

\begin{sol}
Set $n=1919$. Then we have Bernoulli random variables $X_1,\ldots,X_n$
with $X_i=0$ if person $i$ died the week before Passover and $X_i=1$ otherwise.
Let $\theta\in [0,1]$ be the parameter with $X_i\sim \Bern(\theta)$ for each $i$.
So we want to test $H_0:\theta=\theta_0=1/2$ versus $H_1:\theta \neq 1/2$.
We will test this using the Wald test.

Our test statistic is
\[
w = \frac{\hat{\theta} - \theta_0}{\widehat{\se}}.
\]
Our estimate for $\hat{\theta}$ is the mean
\[
\hat{\theta}=\overline{X}=\frac{1}{n}\sum X_i = \frac{997}{1919}\approx 0.5195.
\]
The standard deviation of $\overline{X}$ is $\sigma/\sqrt{n}$ where $\sigma$
is the standard deviation. So we estimate $\se$ by
$\widehat{\se}=\widehat{\sigma}/\sqrt{n}$ where $\widehat{\sigma}$ is the
estimated standard deviation of the $X_i$:
\[
\widehat{\se}^2= \frac{1}{n^2} \sum (X_i-\hat{\theta})^2
= \frac{1}{1919^2} (922 \cdot \hat{\theta}^2 + 997 \cdot (1-\hat{\theta})^2)\approx 0.00013,
\ \ \ \widehat{\se} \approx 0.0114.
\]
Plugging in,
\[
w \approx \frac{0.5195 - 0.5}{0.0114} \approx 1.7105.
\]
By Theorem 10.13, our estimated p-value is
\[
2\Phi(-w)\approx 0.08717.
\]
We can view this as weak evidence against $H_0$ but we don't reject $H_0$
at the size threshold of $0.05$.

A 95\% confidence interval for $\theta$ is given by
\[
(\hat{\theta} - 2 \widehat{\se}, \hat{\theta} - 2\widehat{\se})
= (0.4967, 0.5423).
\]
\end{sol}

\item[(7)] In 1861, 10 essays appeared in the New Orleans Daily Crescent. They
were signed ``Quintus Curtius Snodgras'' and some people suspected
they were actually written by Mark Twain. To investigate this, we will
consider the proportion of three letter words found in an author's work.
From eight Twain essays we have:
\[
.225 \ .262 \ .217 \ .240 \ .230 \ .229 \ .235 \ .217
\]
From 10 Snodgrass essays we have:
\[
.209 \ .205 \ .196 \ .210 \ .202 \ .207 \ .224 \ .223 \ .220 \ .201
\]
\begin{enumerate}[(a)]
\item Perform a Wald test for equality of the means. Use the nonparametric
plug-in estimator. Report the p-value and a 95 per cent confidence
interval for the difference of means. What do you conclude?
\item Now use a permutation test to avoid the use of large sample methods.
What is your conclusion? (Brinegar (1963)).
\end{enumerate}
\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_10/7.ipynb}{7.ipynb}.
\end{sol}

\item[(8)] Let $X_1,\ldots,X_n \sim N(\theta,1)$. Consider testing
\[
H_0 : \theta = 0 \text{ versus } \theta = 1.
\]
Let the rejection region be $R = \{x^n : T(x^n) > c\}$ where
$T(x^n) = n^{-1}\sum_{i=1}^n x_i$.
\begin{enumerate}[(a)]
\item Find $c$ so that the test has size $\alpha$.
\begin{sol}
We have
\[
\sum X_i \sim N(n\theta, n), \ \ \
\overline{X_n} = \frac{1}{n} \sum X_i \sim N(\theta, n/n^2) = N(\theta, 1/n).
\]
So
\[
\beta(\theta)=\P_\theta((X_1,\ldots,X_n) \in R) = \P_\theta(\overline{X_n} > c)
\]
and this is equal to
\[
\P_\theta(\sqrt{n}(\overline{X_n} - \theta) > \sqrt{n}(c-\theta))
= 1 - \Phi(\sqrt{n}(c-\theta)).
\]
The size is
\[
\alpha = \sup_{\theta \in \Theta_0} \beta(\theta) = \beta(0) =
1- \Phi(c\sqrt{n}).
\]
I.e. $\Phi(c\sqrt{n})=1-\alpha$ and solving for $c$ yields
$c=\frac{1}{\sqrt{n}}z_\alpha$.
\end{sol}
\item Find the power under $H_1$, that is, find $\beta(1)$.
\begin{sol}
$\beta(1) = 1-\Phi(\sqrt{n}(c-1))$
\end{sol}
\item Show that $\beta(1) \to 1$ as $n \to \infty$.
\begin{sol}
As $n\to \infty$, $c-1\to -1$ so $\sqrt{n}(c-1)\to -\infty$. Thus
\\ $\Phi(\sqrt{n}(c-1))\to 0$ and $\beta(1)\to 1$.
\end{sol}
\end{enumerate}

\item[(11)]
A randomized, double-blind experiment was conducted to assess the
effectiveness of several drugs for reducing postoperative nausea. The
data are as follows.

\begin{center}
\begin{tabular}{c|c|c}
& Number of Patients & Incidence of Nausea \\
\hline
Placebo & 80 & 45 \\
\hline
Chlorpromazine & 75 & 26 \\
\hline
Dimenhydrinate & 85 & 52 \\
\hline
Pentobarbital (100 mg) & 67 & 35 \\
\hline
Pentobarbital (150 mg) & 85 & 37 \\
\end{tabular}
\end{center}

\begin{enumerate}[(a)]
\item Test each drug versus the placebo at the 5 per cent level. Also, report
the estimated odds-ratios. Summarize your findings.
\item Use the Bonferroni and the FDR method to adjust for multiple testing.
(Beecher (1959)).
\end{enumerate}

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_10/11.ipynb}{11.ipynb}.
\end{sol}

\item[(12)] Let $X_1,\ldots,X_n \sim \Poiss(\lambda)$.
\begin{enumerate}[(a)]
\item Let $\lambda_0>0$. Find the size $\alpha$ Wald test for
\[
H_0:\lambda=\lambda_0 \text{ versus } H_1:\lambda \neq \lambda_0.
\]

\item (Computer Experiment.) Let $\lambda_0=1$, $n = 20$ and $\alpha = .05$.
Simulate $X_1,\ldots,X_n \sim \Poiss(\lambda)$ and perform the Wald test.
Repeat many times and count how often you reject the null.
How close is the type I error rate to $.05$?
\end{enumerate}

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_10/12.ipynb}{12.ipynb}.
\end{sol}

\item[(13)] Let $X_1,\ldots,X_n \sim N(\mu,\sigma^2)$.
Construct the likelihood ratio test for
\[
H_0: \mu=\mu_0 \text{ versus } H_1: \mu \neq \mu_0·
\]
Compare to the Wald test.

\begin{sol}
We assume $\sigma$ to be known.
The likelihood ratio test statistic is
\[
\lambda = 2 \log \left( \frac{\mathcal L(\hat\mu)}{\mathcal L(\mu_0)}\right)
\]
where $\hat \mu$ is the maximum likelihood estimator
$\hat \mu = \overline X = \frac{1}{n}\sum X_i$.

We have
\[
\mathcal L(\hat\mu) = \prod_i \frac{1}{\sqrt{2\pi}\sigma}e^{-(X_i-\hat\mu)^2/2\sigma^2}
= \frac{1}{(\sqrt{2\pi}\sigma)^n}e^{-\sum(X_i-\hat\mu)^2/2\sigma^2}.
\]
Expanding $(X_i-\mu_0)^2$ as
\[
(X_i-\hat\mu)^2+2(\hat\mu-\mu_0)(X_i-\hat\mu) +(\hat\mu-\mu_0)^2
\]
yields
\[
\sum (X_i-\mu_0)^2 = \sum (X_i-\hat \mu)^2 + n(\hat\mu -\mu_0)^2.
\]
So we have
\[
\mathcal L(\mu_0) =
\frac{1}{(\sqrt{2\pi}\sigma)^n} e^{-\left(\sum (X_i-\hat\mu)^2/2\sigma^2 +
n(\hat\mu-\mu_0)^2/2\sigma^2\right)}.
\]
Thus
\[
\frac{\mathcal L(\hat\mu)}{\mathcal L(\mu_0)}=e^{n(\hat\mu-\mu_0)^2/2\sigma^2}
\]
and our statistic is
\[
\lambda = 2 \log \left( \frac{\mathcal L(\hat\mu)}{\mathcal L(\mu_0)}\right)=
n(\hat\mu-\mu_0)^2/\sigma^2.
\]

Say that we reject $H_0$ if $\lambda > c$ for some fixed $c$. This is equivalent
to
\[
\frac{|\hat\mu-\mu_0|}{\sigma/\sqrt{n}} > \sqrt{c}.
\]
Let's compute the size of this test.
Assuming $\mu=\mu_0$, we have that
$\frac{\hat\mu-\mu_0}{\sigma/\sqrt{n}} \sim N(0,1)$.
In this case, the probability of the inequality above is
\[
\Phi(-\sqrt{c}) + (1-\Phi(\sqrt{c})) = 2-2\Phi(\sqrt{c})
\]
and this is the size of our test.
Setting the size to be $\alpha$ yields $c = z_{\alpha/2}^2$. So the size
$\alpha$ likelihood ratio test is
\[
\frac{|\hat\mu-\mu_0|}{\sigma/\sqrt{n}} > z_{\alpha/2}.
\]
The size $\alpha$ Wald test for rejecting $H_0$ is
\[
\frac{|\hat\mu-\mu_0|}{\widehat{\se}} > z_{\alpha/2}.
\]
So we see that the two tests are the same up to replacing $\widehat\se$ (an
estimate of the standard error of $\hat\mu$) by
$\sigma/\sqrt{n}$ (the actual standard error of $\hat\mu$).
\end{sol}


\end{enumerate}
\end{document}
