\documentclass[10pt]{article}

\usepackage{amssymb, amsfonts, amsmath, amsthm}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[margin=1.00 in]{geometry}
\usepackage{enumerate}
\usepackage{harpoon}
\usepackage{nicefrac}
\usepackage{tikz, pgfplots}
\usepackage{xypic}
\usepackage{tikz-cd}
\usepackage{booktabs}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{} % clear all header fields
\renewcommand{\headrulewidth}{0pt} % no line in header area
\fancyfoot{} % clear all footer fields
\fancyfoot[LE,RO]{\footnotesize \thepage}           % page number in "outer" position of footer line

\pgfplotsset{compat=newest}

%\setlength\parindent{0pt}%removes indents from entire file

\newtheorem*{sol}{Solution}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\E}{\mathbb{E}}

\newcommand{\Unif}{\operatorname{Uniform}}
\newcommand{\Bern}{\operatorname{Bernoulli}}
\newcommand{\Binom}{\operatorname{Binomial}}
\newcommand{\Poiss}{\operatorname{Poisson}}
\newcommand{\se}{\operatorname{se}}
\newcommand{\Exp}{\operatorname{Exp}}
\newcommand{\Cov}{\operatorname{Cov}}
\newcommand{\MSE}{\textsc{mse}}
\newcommand{\bias}{\mathsf{bias}}
\newcommand{\CDF}{\textsc{cdf}}

\begin{document}

\noindent \large{Solutions to selected exercises from Chapter 7 of
\emph{Wasserman --- All of Statistics}}

\begin{enumerate}
\item[(1)] Consider the data in Example 8.6. Find the plug-in estimate of the
correlation coefficient. Estimate the standard error using the bootstrap.
Find a 95 percent confidence interval using the normal, pivotal, and
percentile methods.

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_8/1.ipynb}{1.ipynb}.
\end{sol}

\item[(2)] (Computer Experiment.) Conduct a simulation to compare the various
bootstrap confidence interval methods. Let $n = 50$ and let
$T(F)=\int (x-\mu)^3 dF(x)/\sigma^3$ be the skewness. Draw $Y_1,\ldots,Y_n \sim N(0,1)$
and set $X_i=e^{Y_i}, i=1,\ldots,n$. Construct the three types of bootstrap 95
percent intervals for $T(F)$ from the data $X_1,\ldots,X_n$ . Repeat this whole
thing many times and estimate the true coverage of the three intervals.

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_8/2.ipynb}{2.ipynb}.
\end{sol}

\item[(3)] Let
\[
X_1,\ldots,X_n \sim t_3
\]
where $n = 25$. Let $\theta = T(F) = (q_{.75} - q_{.25})/1.34$ where $q_p$ denotes the
$p^{\text{th}}$ quantile. Do a simulation to compare the coverage and length of the
following confidence intervals for $\theta$: (i) normal interval with standard
error from the bootstrap, (ii) bootstrap percentile interval, and (iii)
pivotal bootstrap interval.

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_8/3.ipynb}{3.ipynb}.
\end{sol}

\item[(6)] Computer Experiment.) Let $X_1,\ldots,X_n\sim N(\mu,1$.
Let $\theta=e^\mu$ and let $\hat\theta = e^{\overline X}$.
Create a data set (using $\mu=5$) consisting of $n=100$ observations.
\begin{enumerate}[(a)]
\item Use the bootstrap to get the $\se$ and 95 percent confidence interval
for $\theta$.
\item Plot a histogram of the bootstrap replications. This is an estimate
of the distribution of $\hat\theta$. Compare this to the true sampling distribution
of $\hat\theta$.
\end{enumerate}

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_8/6.ipynb}{6.ipynb}.
\end{sol}

\end{enumerate}
\end{document}
