\documentclass[10pt]{article}

\usepackage{amssymb, amsfonts, amsmath, amsthm}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[margin=1.00 in]{geometry}
\usepackage{enumerate}
\usepackage{harpoon}
\usepackage{nicefrac}
\usepackage{tikz, pgfplots}
\usepackage{xypic}
\usepackage{tikz-cd}
\usepackage{booktabs}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{} % clear all header fields
\renewcommand{\headrulewidth}{0pt} % no line in header area
\fancyfoot{} % clear all footer fields
\fancyfoot[LE,RO]{\footnotesize \thepage}           % page number in "outer" position of footer line

\pgfplotsset{compat=newest}

%\setlength\parindent{0pt}%removes indents from entire file

\newtheorem*{sol}{Solution}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\E}{\mathbb{E}}

\newcommand{\Unif}{\operatorname{Uniform}}
\newcommand{\Bern}{\operatorname{Bernoulli}}
\newcommand{\Binom}{\operatorname{Binomial}}
\newcommand{\Poiss}{\operatorname{Poisson}}
\newcommand{\Gam}{\operatorname{Gamma}}
\newcommand{\se}{\operatorname{se}}
\newcommand{\Exp}{\operatorname{Exp}}
\newcommand{\Cov}{\operatorname{Cov}}
\newcommand{\MSE}{\textsc{mse}}
\newcommand{\bias}{\mathsf{bias}}
\newcommand{\CDF}{\textsc{cdf}}
\newcommand{\MLE}{\textsc{mle}}

\begin{document}

\noindent \large{Solutions to selected exercises from Chapter 9 of
\emph{Wasserman --- All of Statistics}}

\begin{enumerate}
\item[(1)] Let $X_1,\ldots,X_n\sim \Gam(\alpha,\beta)$. Find the method of
moments estimator for $\alpha$ and $\beta$.

\begin{sol}
The PDF is $f(x)=\frac{1}{\beta^\alpha \Gamma(\alpha)}x^{\alpha-1}e^{-x/\beta}$
for $x>0$ and $f(x)=0$ otherwise.
The first moment is $\epsilon_1=\alpha\beta$, as calculated in the text.
The second moment is
\[
\epsilon_2 = \frac{1}{\beta^\alpha \Gamma(\alpha)}
\int_0^\infty x^{\alpha+1e^{-x/\beta}}dx.
\]
Integration by parts yields
\[
\int x^{\alpha+1e^{-x/\beta}}dx = -\beta x^{\alpha+1}e^{-x/\beta} +
\beta(\alpha+1)\int x^{\alpha}e^{-x/\beta}dx
\]
where we note that the first term is zero at $x=0$ and approaches 0 as
$x\to \infty$.
Thus
\[
\epsilon_2 =
\beta(\alpha+1)\int_0^\infty \frac{x^\alpha e^{-x/\beta}}{\beta^\alpha\Gamma(\alpha)}dx
= \beta(\alpha+1)\alpha\beta=\beta^2(\alpha^2+\alpha)
\]
where again we have used the fact that the first moment is $\alpha\beta$.

The sample moments are
\[
\hat \epsilon_1 = \frac{1}{n} \sum X_i = \overline{X_n} \text{ and }
\hat \epsilon_2 = \frac{1}{n} \sum X_i^2.
\]
Our system of equations for the method of moments estimators
$\hat \alpha, \hat \beta$ is
\[
\hat \alpha \hat \beta = \hat \epsilon_1, \ \ \
\hat \beta^2 (\hat \alpha^2 + \hat \alpha) = \hat \epsilon_2.
\]
Setting $\hat \beta=\hat \epsilon_1 / \hat \alpha$, substituting into the
second equation, and solving for $\hat \alpha$ yields
\[
\hat \alpha = \frac{\hat \epsilon_1^2}{\hat \epsilon_2 - \hat \epsilon_1^2}, \ \ \
\hat \beta = \frac{\hat \epsilon_2 - \hat \epsilon_1^2}{\hat \epsilon_1}.
\]
Estimating the mean as $\hat \mu = \hat \epsilon_1$ and the variance as
$\hat \sigma^2 = \hat \epsilon_2 - \hat \epsilon_1^2$, we may rewrite this as
\[
\hat \alpha = \frac{\hat \mu^2}{\hat \sigma^2}, \ \ \
\hat \beta = \frac{\hat \sigma^2}{\hat \mu}.
\]
\end{sol}

\item[(2)]
Let $X_1,\ldots,X_n \sim \Unif(a, b)$ where $a$ and $b$ are unknown parameters
and $a < b$.
\begin{enumerate}[(a)]
\item Find the method of moments estimators for $a$ and $b$.
\item Find the $\MLE$ $\hat a$ and $\hat b$.
\item Let $\tau = \int x dF(x)$. Find the $\MLE$ of $\tau$.
\item Let $\hat\tau$ be the $\MLE$ of $\tau$. Let $\tilde \tau$
be the nonparametric plug-in estimator
of $\tau = \int x dF(x)$. Suppose that $a=1$, $b=3$, and $n=10$. Find the $\MSE$
of $\hat\tau$ by simulation. Find the $\MSE$ of $\tilde\tau$ analytically. Compare.
\end{enumerate}

\begin{sol}
\begin{enumerate}[(a)]
\item The first and second moments are
\[
\alpha_1(a, b) = \frac{1}{2}\frac{b^2-a^2}{b-a} =
\frac{1}{2}(a+b), \ \ \
\alpha_2(a, b) = \frac{1}{3}\frac{b^3-a^3}{b-a} =
\frac{1}{3}(b^2+ab+a^2).
\]
The method of moments estimators satisfy the system
\[
\frac{1}{2}(\hat a + \hat b) = \hat \alpha_1, \ \ \
\frac{1}{3}(\hat a^2 + \hat a \hat b + \hat b^2) = \hat \alpha_2
\]
where $\hat \alpha_1, \hat \alpha_2$ are the sample moments.

Solving the first equation for $\hat b$ yields $\hat b = 2\hat \alpha_1 - \hat a$.
Plugging this into the second equation and simplifying yields
\[
\hat a^2 - 2\hat \alpha_1 \hat a + 4\hat \alpha_1^2 - 3\hat \alpha_2 = 0
\]
which is a degree two polynomial in $\hat a$. Then $\hat a$ is one of the two roots
$\hat \alpha_1 \pm \sqrt{3\hat \alpha_2 - 3 \hat \alpha_1^2}$. Thus
$\hat b = \hat \alpha_1 \mp \sqrt{3\hat \alpha_2 - 3 \hat \alpha_1^2}$. However,
$a<b$ so $\hat b$ must be the larger of the two roots and $\hat a$ must be
the smaller. I.e.
\[
\hat a = \hat \mu - \sqrt{3}\hat \sigma, \ \ \
\hat b = \hat \mu + \sqrt{3}\hat \sigma
\]
where $\hat \mu=\hat \alpha_1$ and $\hat \sigma^2 = \hat \alpha_2 - \hat \alpha_1^2$
are the sample mean and standard deviation, respectively.

\item Set $\theta=(a,b)$ and let $\hat \theta=(\hat a,\hat b)$ be the $\MLE$.
The likelihood of $\theta$ is
\[
\mathcal{L}_n(\theta) = \frac{1}{(b-a)^n}\prod \chi_{[a,b]}(X_i)
\]
where $\chi_\cdot$ denotes an indicator function. This is equal to $0$
if $X_i<a$ or $X_i>b$ for some $i$. Otherwise it is equal to $1/(b-a)^n$.
Thus $\mathcal L_n(\theta)$ is maximized when $b$ is as small as possible, without
being less than any $X_i$ and $a$ is as large as possible, without being larger
than any $X_i$. I.e.
\[
\hat a = \min_i X_i, \ \ \ \hat b = \max_i X_i.
\]
\item We have $\tau = \E(X_i)=(a+b)/2$. By equivariance of the $\MLE$,
the $\MLE$ of $\tau$ is
\[
\hat \tau = \frac{\hat a + \hat b}{2} = \frac{\min X_i + \max X_i}{2}.
\]

\item
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_9/2.ipynb}{2.ipynb}.
\end{enumerate}
\end{sol}

\item[(3)] Let $X_1,\ldots,X_n\sim N(\mu,\sigma^2)$. Let $\tau$ be the .95 percentile,
i.e. $\P(X<\tau) = .95$.
\begin{enumerate}[(a)]
\item Find the $\MLE$ of $\tau$.
\item Find an expression for an approximate $1-\alpha$ confidence interval for
$\tau$.
\item Suppose the data are:
\begin{center}
\begin{tabular}{c c c c c c}
3.23&-2.50&1.88&-0.68&4.43&0.17 \\
1.03&-0.07&-0.01&0.76&1.76&3.18 \\
0.33&-0.31&0.30&-0.61&1.52&5.43 \\
1.54&2.28&0.42&2.33&-1.03&4.00 \\
0.39 &&&&&
\end{tabular}
\end{center}

Find the $\MLE$ $\tau$. Find the standard error using the delta method. Find
the standard error using the parametric bootstrap.
\end{enumerate}

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_9/3.ipynb}{3.ipynb}.
\end{sol}

\item Let $X_1,\ldots,X_n\sim \Poiss(\lambda)$. Find the method of moments estimator,
the maximum likelihood estimator and the Fisher information $I(\lambda)$.

\begin{sol}
For the method of moments, there is only one moment to compute: the first moment.
Thus the sample moment $\overline{X_n}=\frac{1}{n}\sum X_i$ must be equal to
the first moment $\E_{\hat\lambda}(X)=\hat\lambda$. So the method of moments
estimator $\hat\lambda = \overline{X_n}$.

For maximum likelihood, we have $f_{\lambda}(x)=\frac{\lambda^x e^{-\lambda}}{x!}$.
The likelihood of $\lambda$ is
\[
\mathcal L_n(\lambda) =\frac{\lambda^{\sum X_i} e^{-\lambda}}{X_1! \cdots X_n!}
\text{ so }
\log \mathcal L_n(\lambda)= \left(\sum X_i\right)\log \lambda - n\lambda + c
\]
where $c$ is a constant and the derivative with respect to $\lambda$ is
$\sum X_i/\lambda-n$ and this is equal to 0 exactly at
$\lambda=\sum X_i/n = \overline{X_n}$, which is the unique maximum. So
the maximum likelihood estimator $\hat\lambda = \overline{X_n}$.

Finally, the score is
\[
s(x;\lambda)=\frac{\partial \log f(x;\lambda)}{\partial \lambda}
= \frac{\partial}{\partial \lambda} (x\log \lambda - \lambda -\log(x!))
= \frac{x}{\lambda} - 1 \text{ and } \frac{\partial s}{\partial \lambda}
=-\frac{x}{\lambda^2}
\]
The Fisher information is
\[
I(\lambda) = -\E\left(-\frac{X}{\lambda^2}\right) = \lambda / \lambda^2 = 1/\lambda.
\]
\end{sol}

\item[(7)] (Comparing two treatments.) $n_1$ people are given treatment 1 and $n_2$
people are given treatment 2. Let $X_1$ be the number of people on treatment
1 who respond favorably to the treatment and let $X_2$ be the
number of people on treatment 2 who respond favorably. Assume that
$X_1\sim \Binom(n_1,p_1)$ $X_2 \sim \Binom(n_2,p_2)$. Let $\psi=p_1-p_2$
\begin{enumerate}[(a)]
\item Find the $\MLE$ $\hat \psi$ for $\psi$.
\item Find the Fisher information matrix $I(p_1,p_2)$.
\item Use the multiparameter delta method to find the asymptotic standard
error of $\hat \psi$.
\item Suppose that $n_1=n_2=200$, $X_1=160$ and $X_2=148$. Find $\hat \psi$.
Find an approximate 90 percent confidence interval for $\psi$ using (i) the delta
method and (ii) the parametric bootstrap.
\end{enumerate}
\end{enumerate}
\end{document}
