\documentclass[10pt]{article}

\usepackage{amssymb, amsfonts, amsmath, amsthm}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[margin=1.00 in]{geometry}
\usepackage{enumerate}
\usepackage{harpoon}
\usepackage{nicefrac}
\usepackage{tikz, pgfplots}
\usepackage{xypic}
\usepackage{tikz-cd}
\usepackage{booktabs}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{} % clear all header fields
\renewcommand{\headrulewidth}{0pt} % no line in header area
\fancyfoot{} % clear all footer fields
\fancyfoot[LE,RO]{\footnotesize \thepage}           % page number in "outer" position of footer line

\pgfplotsset{compat=newest}

%\setlength\parindent{0pt}%removes indents from entire file

\newtheorem*{sol}{Solution}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\Unif}{\operatorname{Uniform}}
\newcommand{\Bern}{\operatorname{Bernoulli}}
\newcommand{\Binom}{\operatorname{Binomial}}
\newcommand{\Poiss}{\operatorname{Poisson}}
\newcommand{\se}{\operatorname{se}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Cov}{\operatorname{Cov}}
\newcommand{\ol}{\overline}

\begin{document}

\noindent \large{Solutions to selected exercises from Chapter 21 of
\emph{Wasserman --- All of Statistics}}

\begin{enumerate}

\item[(1)] Prove Theorem 21.5.

\begin{sol}
We have $R(J) = \int b(x)^2dx + \int v(x)dx$. Furthermore,
\[
b(x) = \E(\hat f(x)) - f(x) = \sum_{j=1}^J \E(\hat \beta_j) \phi_j(x) - f(x)
= \sum_{j=1}^J \beta_j \phi_j(x) - f(x) = \sum_{j=J+1}^\infty \beta_j \phi_j(x).
\]
By orthogonality of the $\phi_j$,
\[
\int b^2 = \sum_{j=J+1}^\infty \beta_j^2
\]
For the other term,
\[
v(x) = \E\left(\hat f(x) - \E \hat f(x)\right)^2 =
\E \left( \sum_{j=1}^J (\hat \beta_j - \beta_j)\phi_j(x)\right)^2
= \sum_{j=1}^J \sum_{k=1}^J \Cov(\hat \beta_j, \hat \beta_k) \phi_j(x) \phi_k(x).
\]
Again using orthogonality of the $\phi_j$ yields,
\[
\int v(x) dx = \sum_{j=1}^J \V(\hat \beta_j) = \sum_{j=1}^J \frac{\sigma_j^2}{n}.
\]
Adding $\int b^2$ to $\int v$ proves the theorem.
\end{sol}

\item[(4)] Prove Parseval's relation equation (21.6).

\begin{sol}
Set $f_n = \sum_{j=1}^n \beta_j \phi_j$. By hypothesis,
$\int (f-f_n)^2\to 0$ as $n\to \infty$. We have
\[
    \int(f - f_n)^2 = \int f^2 - 2\int f f_n + \int f_n^2.
\]
By the Cauchy-Schwarz inequality (equivalently H\"{o}lder's inequality),
\[
\left|\int f f_n\right| \leq \sqrt{\int f^2} \sqrt{\int f_n^2}.
\]
Hence,
\[
    \int (f - f_n)^2 \geq \int f^2 - 2 \sqrt{\int f^2} \sqrt{\int f_n^2} +
    \int f_n^2 = \left(\sqrt{\int f^2} - \sqrt{\int f_n^2}\right)^2.
\]
Thus $\int f_n^2 \to \int f^2$. On the other hand,
\[
    \int f_n^2 = \sum_{j=1}^n \beta_j^2
\]
by the orthogonality of the $\phi_j$. The proof of H\"{o}lder's inequality
follows from Young's inequality for products. See Wikipedia.
\end{sol}

\item[(5)] Plot the first five Legendre polynomials. Verify, numerically, that they
are orthonormal.

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_21/5.ipynb}{5.ipynb}.
\end{sol}

\item[(6)] Expand the following functions in the cosine basis on $[0, 1]$. For (a)
and (b), find the coefficients $\beta_j$ analytically. For (c) and (d), find the
coefficients $\beta_j$ numerically, i.e.
\[
\beta_j = \int_0^1 f(x)\phi_j(x) dx \approx \frac{1}{N}\sum_{r=1}^N
f\left(\frac{r}{N}\right) \phi_j\left(\frac{r}{N}\right)
\]
for some large integer $N$. Then plot the partial sum $\sum_{j=1}^n\beta_j\phi_j(x)$ for
increasing values of $n$.
\begin{enumerate}[(a)]
\item $f(x) = \sqrt{2}\cos(3\pi x)$
\item $f(x) = \sin(\pi x)$
\item $f(x) = \sum_{j=1}^{11} h_j K(x-t_j)$ where $K(t)=(1+\text{sign}(t))/2$,
$\text{sign}(x)=-1$ if $x<0$, $\text{sign}(x)=0$ if $x=0$, $\text{sign}(x)=1$
$\text{sign}(x)=1$ if $x>0$,

$(t_j) = (.1, .13, .15, .23, .25, .40, .44, .65, .76, .78, .81),$

$(h_j) = (4, -5,3, -4,5, -4.2,2.1,4.3, -3.1,2.1, -4.2)$.
\item $f = \sqrt{x(1-x)}\sin\left(\frac{2.1 \pi}{x + 0.05}\right)$.
\end{enumerate}

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_21/6.ipynb}{6.ipynb}.
\end{sol}

\item[(7)] Consider the glass fragments data from the book's website. Let $Y$ be
refractive index and let $X$ be aluminum content (the fourth variable).
\begin{enumerate}
\item Do a non parametric regression to fit the model $Y = f(x) + \epsilon$ using
the cosine basis method. The data are not on a regular grid. Ignore this
when estimating the function. (But do sort the data first according to
$x$.) Provide a function estimate, an estimate of the risk, and a confidence
band.
\item Use the wavelet method to estimate $f$.
\end{enumerate}

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_21/7.ipynb}{7.ipynb}.
\end{sol}

\item[(9)] Consider again the doppler signal:
\[
f(x) = \sqrt{x(1-x)}\sin\left(\frac{2.1 \pi}{x+0.05}\right).
\]
Let $n=1,024$, $\sigma=0.1$, and let $(x_1,\ldots,x_n)=(1/n,\ldots,1)$. Generate
data
\[
Y_i = f(x_i) + \sigma \epsilon_i
\]
where $\epsilon_i \sim N(0,1)$.
\begin{enumerate}[(a)]
\item Fit the curve using the cosine basis method. Plot the function estimate
and confidence band for $J = 10,20, \ldots,100$.
\item Use Haar wavelets to fit the curve.
\end{enumerate}

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_21/9.ipynb}{9.ipynb}.
\end{sol}

\item[(11)] In this question, we will explore the motivation for equation (21.37). Let
$X_1,\ldots,X_n \sim N(0,\sigma^2)$. Let
\[
\hat \sigma = \sqrt{n} \times \frac{\text{median}(|X_1|,\ldots,|X_n|)}{0.6745}
\]
\begin{enumerate}[(a)]
\item Show that $\E(\hat \sigma) = \sigma$.
\item Simulate $n=100$ observations from a $N(0,1)$ distribution. Compute
$\hat \sigma$ as well as the usual estimate of $\sigma$. Repeat 1,000 times and compare
the MSE.
\item Repeat (b) but add some outliers to the data. To do this, simulate
each observation from a $N(0,1)$ with probability $.95$ and simulate each
observation from a $N(0,10)$ with probability $.95$.
\end{enumerate}

\end{enumerate}
\end{document}
