\documentclass[10pt]{article}

\usepackage{amssymb, amsfonts, amsmath, amsthm}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[margin=1.00 in]{geometry}
\usepackage{enumerate}
\usepackage{harpoon}
\usepackage{nicefrac}
\usepackage{tikz, pgfplots}
\usepackage{xypic}
\usepackage{tikz-cd}
\usepackage{booktabs}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{} % clear all header fields
\renewcommand{\headrulewidth}{0pt} % no line in header area
\fancyfoot{} % clear all footer fields
\fancyfoot[LE,RO]{\footnotesize \thepage}           % page number in "outer" position of footer line

\pgfplotsset{compat=newest}

%\setlength\parindent{0pt}%removes indents from entire file

\newtheorem*{sol}{Solution}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\P}{\mathbb{P}}

\begin{document}

\noindent \large{Solutions to selected exercises from Chapter 1 of
\emph{Wasserman --- All of Statistics}}

\begin{enumerate}
\item[(5)] Suppose we toss a fair coin until we get exactly two heads.
Describe the sample space $S$. What is the probability that exactly
$k$ tosses are required.

\begin{sol}
The sample space is
\[\Omega = \{HH, THH, HTH, HTTH, THTH, TTHH, HTTTH, THTTH, \ldots\}.\]
I.e. an element of $\Omega$ is a sequence in $T$ and $H$
of length $k$ with the last element
being $H$ and exactly one of the first $k-1$ elements being $H$.

There are exactly $k-1$ sequences in $\Omega$ of length $k$, corresponding
to the $k-1$ choices for where to put an $H$ in the initial sequence of
length $k-1$. Each length $k$ sequence has probability $(1/2)^k$ of occurring.
Hence the probability that exactly $k$ tosses are required is $(k-1)(1/2)^k$
\end{sol}

\item[(8)] Suppose that $\P(A_i)=1$ for each $i$. Prove that
\[
\P\left(\bigcap_{i=1}^\infty A_i\right) = 1
\]

\begin{sol}
We have $\P(\bigcap A_i) = 1 -\P(\bigcup A_i^c) \geq 1 - \sum \P(A_i^c)$
where $\cdot^c$ denotes the complement. Since $\P(A_i^c)=0$ for each $i$
we have $\P(\bigcap A_i)=0$, as claimed.
\end{sol}

\item[(11)] Suppose that $A$ and $B$ are independent events. Show that $A^c$
and $B^c$ are independent events.

\begin{sol}
We have $\P(A^cB^c)=\P((A\cup B)^c)=1-\P(A\cup B)=1-\P(A)-\P(B)+\P(AB)$.
Since $A$ and $B$ are independent this is
\[
1-\P(A)-\P(B)+\P(A)\P(B)=(1-\P(A))(1-\P(B)) = \P(A^c)\P(B^c).
\]
\end{sol}

\item[(12)] There are three cards. The first is green on both sides, the
second is red on both sides and third is green on one side and red on the
other. We choose a card at random and we see one side (also chosen at
random). If the side we see is green, what is the probability that the
other side is also green? Many people intuitively answer $1/2$. Show that
the correct answer is $2/3$.

\begin{sol}
The sample space of possible sequences of sides that we can see is
$\Omega = \{RR, GG, RG, GR\}$ where $R$ denotes a red side and $G$ denotes
a green side. The elements of $\Omega$ have probability $1/3$, $1/3$,
$1/6$, and $1/6$. Hence the probability that the other side is green, given
that the first side we see is green is
\[
\frac{1/3}{1/3+1/6} = \frac{2}{3}.
\]
\end{sol}

\item[(13)] Suppose that a fair coin is tossed repeatedly until both a head
and a tail have appeared at least once.
\begin{enumerate}[(a)]
\item Describe the sample space $\Omega$.
\item What is the probability that three tosses will be required?
\end{enumerate}

\begin{sol}
\begin{enumerate}[(a)]
\item The sample space is
\[
\Omega = \{HT, TH, HHT, TTH, HHHT, TTTH, HHHHT, TTTTH, \ldots\}
\]
I.e. an element of $\Omega$ is a sequence of length $n$ where the first
$n-1$ elements are $H$ and the last element is $T$, or vice versa.
In particular, for any $n\geq 2$ there are exactly two sequences in $\Omega$
of length $n$.

\item The probability of three tosses being required is $2(1/2)^3=1/4$ since
there are two elements of $\Omega$ of length $3$, both with probability
$(1/2)^3$.
\end{enumerate}
\end{sol}

\item[(15)] The probability that a child has blue eyes is $1/4$. Assume
independence between children. Consider a family with 3 children.
\begin{enumerate}[(a)]
\item If it is known that at least one child has blue eyes, what is
the probability that at least two children have blue eyes?

\item If it is known that the youngest child has blue eyes, what is
the probability that at least two children have blue eyes?
\end{enumerate}

\begin{sol}
\begin{enumerate}[(a)]
\item Order the children from youngest to oldest. Thus the sample space
consists of sequences of length three in $B$ and $N$, $B$ indicating
that a child has blue eyes and $N$ indicating that they do not have blue
eyes:
\[
\Omega = \{NNN, BNN, NBN, NNB, BBN, BNB, NBB, BBB\}.
\]
If at least one child has blue eyes then the element of the sample space
$\Omega$ can be anything except for $NNN$. The probability of this occuring
is \[
3(1/4)(3/4)^2 + 3(1/4)^2(3/4) + (1/4)^3.
\]
Here the first term is the probability of one of the sequences $BNN,NBN,NNB$,
the second term is the probability of one of the sequences $BBN, BNB, NBB$,
and the third term is the probability of $BBB$. So the probability that
at least two children have blue eyes is
\[3(1/4)^2(3/4) + (1/4)^3.\]
The probability that at least two children have blue eyes given that one child
has blue eyes is
\[
\frac{3(1/4)^2(3/4) + (1/4)^3}{3(1/4)(3/4)^2 + 3(1/4)^2(3/4) + (1/4)^3}
=\frac{10}{37}.
\]
\item If the first child has blue eyes then the sequence is
$BNN, BBN, BNB$, or $BBB$. The probability for this event is
$(1/4)(3/4)^2 + 2(1/4)^2(3/4) + (1/4)^3$. The probability that the first child
has blue eyes and at least two children have blue eyes is
$2(1/4)^2(3/4) + (1/4)^3$. So the conditional probability is
\[
\frac{2(1/4)^2(3/4) + (1/4)^3}{(1/4)(3/4)^2 + 2(1/4)^2(3/4) + (1/4)^3}
=\frac{7}{16}.
\]
\end{enumerate}
\end{sol}

\item[(19)]Suppose that 30 percent of computer owners use a Macintosh, 50 percent
use Windows, and 20 percent use Linux. Suppose that 65 percent of the Mac
users have succumbed to a computer virus, 82 percent of the Windows users get
the virus, and 50 percent of the Linux users get the virus. We select a person
at random and learn that her system was infected with the virus. What is the
probability that she is a Windows user?

\begin{sol}
Let $M$ stand for Mac, $W$ for Windows, $L$ for Linux, and $V$ for virus.
By Bayes's Theorem,
\[
\P(W|V) = \frac{\P(V|W)\P(W)}{\P(V|M)\P(M) +\P(V|W)\P(W) + \P(V|L)\P(L)}.
\]
Substituting yields that $\P(W|V)$ is
\[
\frac{(82/100)(50/100)}{(65/100)(30/100)+(82/100)(50/100)+(50/100)(20/100)}
= \frac{82}{141} \approx 58.156 \%
\]
\end{sol}

\item[(20)] A box contains 5 coins and each has a different probability of
showing heads. Let $p_1,\ldots, p_5$ denotte the probability of heads on each
coin. Suppose that
\[
p_1=0, \ p_2=1/4, \ p_3 =1/2, \ p_4=3/4 \text{ and } p_5=1.
\]
Let $H$ denote ``heads is obtained'' and let $C_i$ denote the event that
coin $i$ is selected.
\begin{enumerate}[(a)]
\item Select a coin at random and toss it. Suppose a head is obtained.
What is the posterior probability that coin $i$ was selected
($i = 1, ... , 5$)?
In other words, find $\P(C_i|H)$ for $i = 1, ... ,5$.
\item Toss the coin again. What is the probability of another head? In
other words find $\P(H_2|H_1)$ where $H_j$ $=$ ``heads on toss $j$.''

Now suppose that the experiment was carried out as follows: We select
a coin at random and toss it until a head is obtained.
\item Find $\P(C_i|B_4)$ where $B_4$ $=$ ``first head is obtained
on toss 4.''
\end{enumerate}

\begin{sol}
\begin{enumerate}[(a)]
\item By Bayes's Theorem:
\[
\P(C_i | H) = \frac{\P(H|C_i)\P(C_i)}{\sum_j \P(H | C_j)\P(C_j)}=
\frac{p_i(1/5)}{\sum_j p_j(1/5)}=\frac{p_i}{1/4+1/2+3/4+1} = \frac{2p_i}{5}.
\]
I.e. the probabilities are
\[
0, \ 1/10, \ 1/5, \ 3/10, \ 2/5.
\]
\item We have $\P(H_2|H_1)=\P(H_1H_2)/\P(H_1)$ and $\P(H_1)=\sum p_i/5 =1/2$.
We also have
\[
\P(H_1H_2) = \sum_{i=1}^5 \P(H_1H_2C_i) = \sum_i \P(H_1H_2|C_i)\P(C_i)
=1/5 \sum_i p_i^2 =3/8.
\]
Therefore
\[
\P(H_2|H_1) = \frac{3/8}{1/2} = \frac{3}{4}.
\]
\item By Bayes's Theorem:
\[
\P(C_i|B_4)= \frac{\P(B_4|C_i)\P(C_i)}{\sum_j \P(B_4|C_j)\P(C_j)}
=\frac{(1-p_i)^3p_i(1/5)}{\sum_j(1-p_j)^3p_j(1/5)} =
\frac{128(1-p_i)^3p_i}{23}.
\]
The probabilities are
\[
0, \ 27/46, \ 8/23, \ 3/46, \ 0.
\]
\end{enumerate}
\end{sol}

\item[(22)] (Computer Experiment.) Suppose we flip a coin $n$ times and let $p$
denote the probability of heads. Let $X$ be the number of heads. We call $X$
a binomial random variable, which is discussed in the next chapter.
Intuition suggests that $X$ will be close to $np$. To see if this is true, we
can repeat this experiment many times and average the $X$ values. Carry
out a simulation and compare the average of the X's to $np$. Try this for
$p =.3$ and $n = 10$, $n = 100$, and $n = 1,000$.

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_1/22.ipynb}{22.ipynb}.
\end{sol}

\item[(23)] (Computer Experiment.) Here we will get some experience simulating
conditional probabilities. Consider tossing a fair die. Let $A = \{2, 4, 6\}$
and $B = \{1, 2, 3, 4\}$. Then, $\P(A) = 1/2$, $\P(B) = 2/3$ and
$\P(AB) = 1/3$. Since $\P(AB) = \P(A)\P(B)$, the events $A$ and $B$ are
independent. Simulate draws from the sample space and verify that
$\hat{\P}(AB) = \hat{\P}(A)\hat{\P}(B)$ where $\hat{\P}(A)$
is the proportion of times $A$ occurred in the simulation and
similarly for $\hat{\P}(AB)$ and $\hat{\P}(B)$. Now find two events
$A$ and $B$ that are not independent. Compute $\hat{\P}(A)$, $\hat{\P}(B)$
and $\hat{\P}(AB)$. Compare the calculated
values to their theoretical values. Report your results and interpret.
\end{enumerate}

\begin{sol}
See the Jupyter Notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_1/23.ipynb}{23.ipynb}.
\end{sol}
\end{document}
