\documentclass[10pt]{article}

\usepackage{amssymb, amsfonts, amsmath, amsthm}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[margin=1.00 in]{geometry}
\usepackage{enumerate}
\usepackage{harpoon}
\usepackage{nicefrac}
\usepackage{tikz, pgfplots}
\usepackage{xypic}
\usepackage{tikz-cd}
\usepackage{booktabs}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{} % clear all header fields
\renewcommand{\headrulewidth}{0pt} % no line in header area
\fancyfoot{} % clear all footer fields
\fancyfoot[LE,RO]{\footnotesize \thepage}           % page number in "outer" position of footer line

\pgfplotsset{compat=newest}

%\setlength\parindent{0pt}%removes indents from entire file

\newtheorem*{sol}{Solution}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\E}{\mathbb{E}}

\newcommand{\Unif}{\operatorname{Uniform}}
\newcommand{\Bern}{\operatorname{Bernoulli}}
\newcommand{\Binom}{\operatorname{Binomial}}
\newcommand{\Poiss}{\operatorname{Poisson}}
\newcommand{\se}{\operatorname{se}}
\newcommand{\Exp}{\operatorname{Exp}}
\newcommand{\Cov}{\operatorname{Cov}}

\begin{document}

\noindent \large{Solutions to selected exercises from Chapter 3 of
\emph{Wasserman --- All of Statistics}}

\begin{enumerate}
\item[(1)]
Suppose we play a game where we start with $c$ dollars. On each play of
the game you either double or halve your money, with equal probability.
What is your expected fortune after $n$ trials?

\begin{sol}
Suppose you halve your money $m$ times out of $n$. There are $\binom{n}{m}$
ways to do this, each with probability $(1/2)^n$. Your fortune after $n$
trials is $(1/2)^m2^nc$. We get the expected value of your fortune after $n$
trials by summing over $m$:
\[
\sum_{m=0}^n \binom{n}{m}\left(\frac{1}{2}\right)^n \left(\frac{1}{2}\right)^m2^{n-m}c
= \sum_{m=0}^n \binom{n}{m} \left(\frac{1}{2}\right)^{2m}c.
\]
We can recognize this as
\[
\sum_{m=0}^n \binom{n}{m} \left(\frac{1}{4}\right)^m(1)^{n-m} =
c\left(1+\frac{1}{4}\right)^n = c\left(\frac{5}{4}\right)^n.
\]
\end{sol}

\item[(3)]
Let $X_1, \ldots ,X_n \sim \Unif(0,1)$ and let $Y_n = \max\{X_1, ... ,X_n\}$. Find
$\E(Y_n)$.

\begin{sol}
We suppose the $X_i$ are independent. For $x\in [0,1]$, the CDF of $Y$ satisfies
\[
F_Y(x)=\P(X_i\leq x \text{ for all } i)=x^n
\]
and $f_Y(x) = nx^{n-1}$ on $[0,1]$ and $0$ elsewhere.

Thus
\[
\E(Y) = \int_0^1 x (n x^{n-1}) dx =\frac{n}{n+1}.
\]
\end{sol}

\item[(4)]
A particle starts at the origin of the real line and moves along the line in
jumps of one unit. For each jump the probability is $p$ that the particle
will jump one unit to the left and the probability is $1-p$ that the particle
will jump one unit to the right. Let $X_n$ be the position of the particle
after $n$ units. Find $\E(X_n)$ and $\V(X_n)$. (This is known as a \textbf{random
walk}.)

\begin{sol}
Let $Y_1,\ldots,Y_n$ be independent $\Bern(p)$ random variables with $Y_i=0$
if the particle moves to the right on the $i^{\text{th}}$ step and $Y_i=1$
otherwise. Then
\[
X_n = \sum_{i=1}^n (1-2Y_i).
\]
We have $\E(X_n) = \sum_{i=0}^n \E(1-2Y_i)$ and
$\E(1-2Y_i) = (1-p)(1) + p(-1)=1-2p$ so $\E(X_n)=n(1-2p)$.

By independence, the variance is $\V(X_n) = \sum_{i=0}^n \V(1-2Y_i)$.
We have $(1-2Y_i)^2=1$ always so $\E(((1-2Y_i)^2)=1$. So
\[
\V(1-2Y_i) = \E((1-2Y_i)^2) - \E(1-2Y_i)^2 = 1 - (1-2p)^2 = 4p-4p^2
\text{ and } \V(X_n) = 4np(1-p).
\]
\end{sol}

\item[(5)]
A fair coin is tossed until a head is obtained. What is the expected
number of tosses that will be required?

\begin{sol}
The sample space is $\{H, TH, TTH, TTTH,\ldots\}$ with the $i^{th}$ entry
occurring with probability $1/2^i$. The expected value is thus
\[
\frac{1}{2}(1) + \frac{1}{4}(2) + \frac{1}{8}(3) + \ldots =
\sum_{i=0}^\infty \frac{i}{2^i}.
\]
Recall that the Taylor series of $1/(1-x^2)$ is $\sum i x^{i-1}$. So
the expected value is
\[
\frac{1}{2}\sum_{i=0}^\infty \frac{i}{2^{i-1}} =
\frac{1}{2} \frac{1}{(1-\frac{1}{2})^2} = 2.
\]
\end{sol}

\item[(9)]
(Computer Experiment.) Let $X_1, X_2, \ldots, X_n$ be $N(0, 1)$ random variables
and let $\overline{X_n} = n^{-1} \sum_{i=1}^n X_i$.
Plot $X_n$ versus $n$ for $n = 1, \ldots, 10,000$.
Repeat for $X_1,X_2,\ldots,X_n \sim \operatorname{Cauchy}$.
Explain why there is such a difference.

\begin{sol}
See the Jupyter notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_3/9.ipynb}{9.ipynb}.
\end{sol}

\item[(10)]
Let $X \sim N(0,1)$ and let $Y = e^X$. Find $\E(Y)$ and $\V(Y)$.

\begin{sol}
By the rule of the lazy statistician,
\[
\E(Y) = \frac{1}{\sqrt{2\pi}} \int e^x e^{-x^2/2}dx
= \frac{1}{\sqrt{2\pi}} \int e^{-(x-1)^2/2 + 1/2}dx
= \sqrt{e} \int \frac{1}{\sqrt{2\pi}} e^{-(x-1)^2/2}dx.
\]
The integral $\int \frac{1}{\sqrt{2\pi}} e^{-(x-1)^2/2}dx$ is just the total
integral of the PDF of a $N(1,1)$ random variable so it's equal to 1. Finally
then $\E(Y)=\sqrt{e}$.

To compute the variance we consider $Y^2 = e^{2X}$. We have
\[
\E(Y^2) = \int \frac{1}{\sqrt{2\pi}} e^{1/2(4x-x^2)}dx =
\int \frac{1}{\sqrt{2\pi}} e^{-1/2(2-x)^2+2}dx =
e^2 \int \frac{1}{\sqrt{2\pi}} e^{-1/2(2-x)^2}dx = e^2.
\]

Finally then
\[
\V(Y) = \E(Y^2) - \E(Y)^2 = e^2-e.
\]
\end{sol}

\item[(11)]
(Computer Experiment: Simulating the Stock Market.) Let $Y_1,Y_2,\ldots$ be
independent random variables such that $\P(Y_i = 1) = \P(Y_i = -1) =
1/2$. Let $X_n = \sum_{i=1}^n Y_i$. Think of $Y_i=1$ as
``the stock price increased by one dollar'', $Y_i=-1$ as
``the stock price decreased by one dollar'',
and $X_n$ as the value of the stock on day $n$.
\begin{enumerate}[(a)]
\item Find $\E(X_n)$ and $\V(X_n)$.
\item Simulate $X_n$ and plot $X_n$ versus $n$ for $n = 1,2, \ldots, 10,000$.
Repeat the whole simulation several times. Notice two things. First, it's easy
to ``see'' patterns in the sequence even though it is random. Second,
you will find that the four runs look very different even though they
were generated the same way. How do the calculations in (a) explain
the second observation?
\end{enumerate}

\begin{sol}
\begin{enumerate}[(a)]
\item
First
\[
\E(X_n) = \sum \E(X_i) = 0.
\]
We have $\V(X_i)=\frac{1}{2}(1-0)^2 + \frac{1}{2}(-1-0)^2=1$ so
\[
\V(X_n) = \sum \V(X_i) = n.
\]
\item See the Jupyter notebook
\href{https://github.com/ajrasmus/some_of_statistics/blob/main/chapter_3/11.ipynb}{11.ipynb}.
\end{enumerate}
\end{sol}

\item[(15)]
Let
\[
f_{X,Y}(x,y) =
\begin{cases}
\frac{1}{3}(x+y) & 0\leq x \leq 1, 0\leq y \leq 1 \\
0 & \text{otherwise}
\end{cases}
\]
Find $\V(2X - 3Y + 8)$.

\begin{sol}
Set $Z=r(X,Y)=2X-3Y$. We have $\V(2X-3Y+8)=\V(Z)=\E(Z^2) - \E(Z)^2$.
We'll calculate these two terms separately.
\[
\E(Z) = \int_0^2 \int_0^1(2x-3y) \frac{1}{3}(x+y)dxdy=-\frac{23}{9}.
\]
And $Z^2=4X^2-12XY+9Y^2$ so
\[
\E(Z^2)=\int_0^2\int_0^1 (4x^2-12xy+9y^2)\frac{1}{3}(x+y)dxdy=\frac{86}{9}.
\]
So
\[
\V(Z)= \frac{86}{9} - \left(\frac{23}{9}\right)^2 = \frac{245}{81}.
\]
\end{sol}

\item[(21)]
Let $X$ and $Y$ be random variables. Suppose that $\E(Y|X) = X$. Show
that $\Cov(X, Y) = V(X)$.

\begin{sol}
By hypothesis,
\[
x = \E(Y | X=x) = \int y \frac{f_{X,Y}(x,y)}{f_X(x)}dy
\]
i.e. $\int y f_{X,Y}(x,y)dy = x f_X(x)$.

Letting $\mu_X,\mu_Y$ be the expectations, we have
\[
\Cov(X,Y) = \E((X-\mu_X)(Y-\mu_Y)) =
\int \int (x-\mu_X)(y-\mu_Y)f_{X,Y}(x,y)dxdy
\]
and this is equal to
\[
\int (x-\mu_X)\left(\int (y-\mu_Y) f_{X,Y}(x,y)dy\right)dx
= \int (x-\mu_X) \left(xf_X(x) - \mu_Yf_X(x)\right)dx
\]
using that $\int y f_{X,Y}(x,y)dy = x f_X(x)$.

Now note by the Rule of Iterated Expectations that
\[
\mu_Y = \E(Y) = \E(\E(Y|X)) = \E(X) = \mu_X.
\]
plugging this into the equality from the last paragraph gives
\[
\Cov(X,Y) = \int (x-\mu_X)(x-\mu_Y)f_X(x)dx = \int(x-\mu_X)^2 f_X(x)dx=\V(X).
\]
\end{sol}

\item[(22)]
Let $X \sim \Unif(0,1)$. Let $0 < a < b < 1$. Let
\[
Y =
\begin{cases}
1 & 0<x<b \\
0 & \text{otherwise}
\end{cases}
\]
and let
\[Z =
\begin{cases}
1 & a < x < 1 \\
0 & \text{otherwise}
\end{cases}
\]
\begin{enumerate}[(a)]
\item Are $Y$ and $Z$ independent? Why/Why not?
\item Find $\E(Y|Z)$. Hint: What values $z$ can $Z$ take?
Now find $\E(Y|Z=z)$.
\end{enumerate}

\begin{sol}
\begin{enumerate}[(a)]
\item $Y$ and $Z$ are not independent. For $\P(Y=1,Z=1)=b-a$ whereas
$\P(Y=1)\P(Z=1) = b(1-a) > b-a$.
\end{enumerate}
\end{sol}
\end{enumerate}
\end{document}
